Outline paper - Multiple imputation for statistical disclosure control: Creating synthetic datasets with *mice*

### 2020-09-10 mice_synthesizing/paper.Rmd
update GV

Open science, including open data, has been marked as the future of science [@gewin_data_2016], and the advantages of publicly available research data are numerous [@molloy_open_2011; @walport_brest_sharing_2011]. As collecting research data requires an enormous investment both in terms of time and monetary resources, openly accessible research data bears the potential of increasing the scientific returns. Additionally, the fact that public funds are used for data collection results in increased demands for the collected data. Nevertheless, the possibilities to share research data directly are often very limited due to restrictions with regard to privacy and confidentiality. Simply anonimizing the data is not enough to fulfil these requirements [@ohm_broken_2009]. Over the years, several other techniques have been used to increase the confidentiality of the data, such as categorizing continuous variables, top coding values above a certain threshold or adding random noise [@drechsler_synthetic_2011]. However, these methods may distort relationships between variables, reducing the data quality drastically.

An alternative solution has been proposed separately by @rubin_statistical_disclosure_1993 and @little_statistical_1993. Although their approaches differed to some extent, the general concept was to release multiply imputed synthetic datasets, instead of the actually observed data. Based on some imputation model, values from the actually observed data are replaced by multiple draws from the posterior predictive distribution of the observed data. Using this approach, the researcher could replace the observed dataset as a whole with multiple synthetic versions. Additionally, the researcher could opt to only replace a subset of the observed data. The values that are to be replaced can be complete variables that could be compared with publicly available datasets or registers, such as addresses, or can be certain values that bear a high risk of being disclosive, say, income values above a certain threshold.

Conceptually, the synthetic data framework is based upon the building blocks of multiple imputation for missing data, as proposed by @rubin_multiple_1987. Now, instead of replacing the missing values with multiple draws from the posterior predictive distribution, the sensitive values are replaced accordingly. Similarly to analysing multiply imputed datasets dealing with missing data, multiply imputed synthetic datasets can be analysed using conventional analysing techniques. The results of the analyses on each synthetic dataset should be pooled, so that the researcher obtains the estimates of the quantity of interest, as well as the estimate of its variance that adequately account for the additional variability that is induced by the imputation procedure.

Potentially, this approach could fulfil the needs for openly accessibly data, without running into barriers with regard to privacy and confidentiality constraints. However, there is no such thing as a free lunch: data collectors have to put effort in creating high-quality synthetic data. Also, the quality of the synthetic data are highly dependent on the imputation models, and using flawed models to generate synthetic data might bias subsequent analyses. Conversely, if the models used to create the synthetic data are able to preserve the relationships between the variables as in the original data, the synthetic data can be nearly as informative as the observed data. Thus, to fully exploit the benefits of synthetic data, the effort to actually create these datasets should be kept at a minimum.

>>>>

Open science, including open data, has been marked as the future of science [@gewin_data_2016], and the advantages of publicly available research data are numerous [@molloy_open_2011; @walport_brest_sharing_2011]. Because collecting research data requires an enormous investment both in terms of time and monetary resources, openly accessible research data bears the potential of increasing the scientific returns. Additionally, the fact that public funds are used for data collection results in increased demands for the collected data. Nevertheless, the possibilities to share research data directly are often very limited due to restrictions with regard to privacy and confidentiality. Simply anonimizing the data is not enough to fulfill these requirements [@ohm_broken_2009]. Over the years, several other techniques have been used to increase the confidentiality of the data, such as categorizing continuous variables, top coding values above a certain threshold or adding random noise [@drechsler_synthetic_2011]. However, these methods may distort relationships between variables, thereby drastically reducing the data quality and the usability for further research.

An alternative solution has been proposed separately by @rubin_statistical_disclosure_1993 and @little_statistical_1993. Although their approaches differed to some extent, the general thought is to use actually observed data to generated multiply imputed synthetic datasets that can be freely disclosed. Based on some imputation model, values from the actually observed data are replaced by multiple draws from the posterior predictive distribution of the observed data. Using this approach, the researcher could replace the observed data set as a whole with multiple synthetic versions. Additionally, the researcher could opt to only replace a subset of the observed data. The values that are to be replaced can be complete variables that could be compared with publicly available data sets or registers, such as addresses, or can be certain values that bear a high risk of being disclosive, say, income values above a certain threshold.

Conceptually, the synthetic data framework is based upon the building blocks of multiple imputation for missing data, as proposed by @rubin_multiple_1987. Instead of replacing just the missing values with multiple draws from the posterior predictive distribution, one could easily *overimpute* any observed sensitive values. Doing this multiple times would allow for drawing correct statistical inferences, similar to those obtained from ordinary multiple imputation for incomplete data sets. The results of the multiply synthetisized data sets should be pooled into an single inference, so that the researcher can draw valid inference from a single pooled analysis. To that respect, the variance should reflect the added variability that is induced by the imputation procedure.

Potentially, this approach could fulfill the needs for openly accessibly data, without running into barriers with regard to privacy and confidentiality constraints. However, there is no such thing as a free lunch: data collectors have to put effort in creating high-quality synthetic data. Also, the quality of the synthetic data are highly dependent on the imputation models, and using flawed models to generate synthetic data might bias subsequent analyses. Conversely, if the models used to create the synthetic data are able to preserve the relationships between the variables as in the original data, the synthetic data can be nearly as informative as the observed data. Thus, to fully exploit the benefits of synthetic data, the effort to actually create these datasets should be kept at a minimum.
